---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# isaeditor

<!-- badges: start -->
<!-- badges: start -->
[![R-CMD-check](https://github.com/bihealth/isaeditor/workflows/R-CMD-check/badge.svg)](https://github.com/bihealth/isaeditor/actions)
<!-- badges: end -->
<!-- badges: end -->

isaeditor is a collection of helper functions for modifying and displaying
[ISA-Tab](https://isa-tools.org/) files. 

## Installation

You can install the released version of isaeditor from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("isaeditor")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bihealth/isaeditor")
```


## Basic usage

### Reading ISA-Tabs

Use the `read_isa()` function:

```{r}
library(isaeditor)
file  <- system.file('extdata', 's_isatab.txt', package='isaeditor')
isa_s <- read_isa(file)
dim(isa_s)
class(isa_s)
summary(isa_s)
print(isa_s)
```

You can directly modify the isatab object almost as simply as you would do
it with a data frame:

```{r}
## access a node
isa_s[ "New Node" ] <- c("em", "pstrem", "bzdrem")

## create a property of the new node
isa_s[ "New Node", "Characteristics[Replicate]" ] <- 1:3

## remove the node and all its properties
isa_s[ "New Node" ] <- NULL
```

Unfortunately, multiple nodes with the same label may exist according to
the ISA-Tab specifications. Sometimes it is therefore necessary to indicate
which of these nodes we mean. There are several ways to do it in
`isaeditor`, two of them are shown here:

```{r}
file <- system.file('extdata', 'a_isatab.txt', package='isaeditor')
isa_a <- read_isa(file)

## use the internal ID to access the node
## you can also use isa_ID_find for that
isa_nodes(isa_a)
isa_a[['ID34']]

## specify which of the nodes 
isa_a[ "Extract Name", n=2 ]
```



